Operational adoption of new probabilistic point rainfall forecasts: the crucial role of user-centric training

### 2020-11-27 manuscript/main.tex

Numerical Weather Prediction (NWP) models produce forecasts that, although improve continuously in reliability and skill, represent an average over the model grid-box. Whilst, for some applications, users require forecasts at specific locations. Despite NWP modelling advancements, significant weather events will still happen at a location within the gaps, in time and space, not picked by the models. Experienced forecasters add value to predictions issued for smaller scales than the model resolution as they understand model biases and model errors at (sub-)grid scale, interpret how they might impact the weather at a specific location, and summarize all this information in an intelligible way for the relevant users. 
However, forecasters call upon past model performance experience, which can take years to build up, might generate biased forecasts based on what such experience can recall, and can take some time to adapt it to new model versions.

@@

Section 2 discusses the experience of Hungary and Costa Rica in probabilistic weather forecasting. Section 3  introduces ecPoint-Rainfall. Section 4 presents the design of products based on ecPoint-Rainfall developed at both meteorological services and how they were subsequently used in a subjective evaluation of the post-processed rainfall forecasts. Section 5 discusses the outcomes of such subjective evaluation. Concluding remarks are presented in section 5.

@@

\begin{figure}
\centerline{\includegraphics[width=19pc]{manuscript/Figures/Fig2.jpg}}
\caption{Orography of Costa Rica (a) and Hungary (b). The insert box shows the location of Hungary and Costa Rica in Europe and America, respectively.}
\label{Fig2}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=18pc]{manuscript/Figures/Fig5.jpg}}
\caption{Product developed by IMN based on the 85th percentile of ecPoint-Rainfall.}
\label{Fig5}
\end{figure}

During the trial, IMN had access to the 99 percentiles for ecPoint-Rainfall. However, IMN developed a product based only on the 85th percentile of ecPoint-Rainfall (see Fig.\ref{Fig5}). The selection of the 85th percentile was not based on previous experience with probabilistic products \textit{("en realidad no solemos usar muy seguido herramientas probabilísticas para pronóstico} del tiempo."} [translation: actually, we don't often use probabilistic tools to forecast the weather.]). The selection of the 85th percentile was not carried out through a formal study. IMN reported that the 85th percentile (and not, for example, the 99th) was selected in order to forecast also more common events that, even if they are no that extreme, can still cause severe impacts in certain regions of Costa Rica.

@@

%%%%%%%%%%%%%%%%%%%%
% Methods
\section{Methods}

\begin{figure*}
\centerline{\includegraphics[width=39pc]{manuscript/Figures/Fig3.png}}
\caption{Timeline of the trial phases. The first phase, the "real-time phase", took place between 2018 and 2019 and lasted 12 months. Th second phase, the "offline phase", took place between 2019 and 2021 and lasted 15 months.}
\label{Fig3}
\end{figure*}

\subsection{Hands-on experience, the trial}
 OMSZ and IMN volunteered to test the ecPoint-Rainfall forecasts in their day-to-day operational routines. The participation of OMSZ and IMN was ideal as the two meteorological services have different degrees of experience with probabilistic forecasts, and reside in two different climatological regions; tropical for IMN and extra-tropical for OMSZ. The trial period started in 2018. During the trial, OMSZ and IMN were invited to analyse the ecPoint-Rainfall performance in predicting extreme rainfall events. Given the small volumes of data available (only one year of data), both countries decided to base their analysis case studies. The selection of the case studies . The ecPoint-Rainfall forecasts were presented to both countries at the beginning of the trial. They were informed of different sources of information about ecPoint-Rainfall on the web, such as newsletters or the ECMWF Forecaster's User Guide \citep{Owens2018} and no further details were provided unless direct requests were submitted.

>>>>

Numerical Weather Prediction (NWP) models have been continuously improving over the last years,  although improve continuously in reliability and skill, represent an average over the model grid-box.

Whilst, for some applications, users require forecasts at specific locations. Despite NWP modelling advancements, significant weather events will still happen at a location within the gaps, in time and space, not picked by the models. Experienced forecasters add value to predictions issued for smaller scales than the model resolution as they understand model biases and model errors at (sub-)grid scale, interpret how they might impact the weather at a specific location, and summarize all this information in an intelligible way for the relevant users. 
However, forecasters call upon past model performance experience, which can take years to build up, might generate biased forecasts based on what such experience can recall, and can take some time to adapt it to new model versions.

@@

Section 2 discusses the experience of Hungary and Costa Rica in probabilistic weather forecasting. Section 3  introduces ecPoint-Rainfall. Section 4 presents the design of products based on ecPoint-Rainfall developed at both meteorological services and how they were subsequently used in a  evaluation of the post-processed rainfall forecasts. Section 5 discusses the outcomes of such subjective evaluation. Concluding remarks are presented in section 5.

@@

\begin{figure}
\centerline{\includegraphics[width=19pc]{manuscript/Figures/Fig2.png}}
\caption{Orography of Costa Rica (a) and Hungary (b). The insert box shows the location of Hungary and Costa Rica in Europe and America, respectively.}
\label{Fig2}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=18pc]{manuscript/Figures/Fig3.jpg}}
\caption{Product developed by IMN based on the 85th percentile of ecPoint-Rainfall.}
\label{Fig3}
\end{figure}

During the trial, IMN had access to the 99 percentiles for ecPoint-Rainfall. However, IMN developed a product based only on the 85th percentile of ecPoint-Rainfall (see Fig.\ref{Fig3}). The selection of the 85th percentile was not based on previous experience with probabilistic products \textit{("en realidad no solemos usar muy seguido herramientas probabilísticas para pronóstico} del tiempo."} [translation: actually, we don't often use probabilistic tools to forecast the weather.]). The selection of the 85th percentile was not carried out through a formal study. IMN reported that the 85th percentile (and not, for example, the 99th) was selected in order to forecast also more common events that, even if they are no that extreme, can still cause severe impacts in certain regions of Costa Rica.

@@

%%%%%%%%%%%%%%%%%%%%
% Methods
\section{The trial}

\begin{figure*}
\centerline{\includegraphics[width=39pc]{manuscript/Figures/Fig4.png}}
\caption{Timeline of the trial phases. The first phase, the "real-time phase", took place between 2018 and 2019 and lasted 12 months. Th second phase, the "offline phase", took place between 2019 and 2021 and lasted 15 months.}
\label{Fig4}
\end{figure*}

In 2018, OMSZ and IMN volunteered to evaluate ecPoint-Rainfall forecasts in the prediction of extreme localized rainfall events in their countries in their day-to-day operational routines. The participation of OMSZ and IMN was ideal. The two meteorological services have different degrees of experience with probabilistic forecasts. IMN has much more recent experience than OMSZ working with PEFs (as mentioned in sections 3b and 4b). The two countries also reside in two different climatological regions. Costa Rica is located in the tropic and Hungary in the extra-tropics (see Fig.\ref{Fig2}a). This would allow to evaluate ecPoint-Rainfall in for different types of rainfall events and two different geographical regions.

\subsection{First phase: the real-time phase}
 \begin{figure}
\centerline{\includegraphics[width=19pc]{manuscript/Figures/Fig5.png}}
\caption{Timeline of the trial phases. The first phase, the "real-time phase", took place between 2018 and 2019 and lasted 12 months. Th second phase, the "offline phase", took place between 2019 and 2021 and lasted 15 months.}
\label{Fig4}
\end{figure}

 The trial started with a one month "Remote training" phase. The materials provided to IMN and OMSZ are listed in Table 1.

 depending on the meteorological service experience with PEFs, was more involving for Costa Rica. The raining material

 Given the small volumes of data available (only one year of data), both countries decided to base their analysis case studies. The selection of the case studies . The ecPoint-Rainfall forecasts were presented to both countries at the beginning of the trial. They were informed of different sources of information about ecPoint-Rainfall on the web, such as newsletters or the ECMWF Forecaster's User Guide \citep{Owens2018} and no further details were provided unless direct requests were submitted.
